---
title: "Linear Regression"
output: html_notebook
---

# Read the train and test files

```{r}
pisaTrain<-read.csv("pisa2009train.csv")
pisaTest<-read.csv("pisa2009test.csv")
```

# Find the Reading Score by Gender

```{r}
with(pisaTrain,tapply(readingScore,male,mean,na.rm=TRUE))
```

# Remove missing values
```{r}
pisaTrain = na.omit(pisaTrain)
pisaTest = na.omit(pisaTest)  
```

Given "White" is the predominant race in the US, let us use that as reference.
Remember that for categorical independent variables, coeffients of other levels are evaluated with respect to a base. Usually the first categorical value is chosen as a base.

```{r}
pisaTrain$raceeth = relevel(pisaTrain$raceeth, "White")
pisaTest$raceeth = relevel(pisaTest$raceeth, "White")
```

# Convert the male variable (indicating gender) to factor for better interpretability
```{r}
pisaTrain$male = as.factor(pisaTrain$male)
pisaTest$male = as.factor(pisaTest$male)
```

# Run the model
```{r}
lmScore<-lm(readingScore~.,data=pisaTrain)
summary(lmScore)
```

# Compute training RMSE
```{r}
SSE = sum(lmScore$residuals^2)
SSE
RMSE = sqrt(SSE/nrow(pisaTrain))
RMSE
```

# Generate Predictions
```{r}
predTest<-predict(lmScore,newdata=pisaTest)
summary(predTest)
```

# calculate test RMSE and R2
```{r}
SSE = sum((pisaTest$readingScore - predTest)^2)
RMSE = sqrt(SSE/nrow(pisaTest))
mean(pisaTest$readingScore)
SST=sum((pisaTest$readingScore - mean(pisaTest$readingScore))^2)
R2=1-(SSE/SST)
```

